---
title: "Quiz2"
author: "Shubham Gupta"
date: "August 13, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1
```{r Q1}
library(gplots)
library(devtools)
library(Biobase)
library(broom)
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/montpick_eset.RData")
load(file=con)
close(con)
mp = montpick.eset
pdata=pData(mp)
edata=as.data.frame(exprs(mp))
fdata = fData(mp)
edata_tr = log2(edata + 1)
svd1 = svd(edata)
svd2 = svd(edata_tr)
names(svd1)
svd1$d[1]^2/sum(svd1$d^2)
svd2$d[1]^2/sum(svd2$d^2)
edata_centered = edata_tr - rowMeans(edata_tr)
svd3 = svd(edata_centered)
svd3$d[1]^2/sum(svd3$d^2)
```

### Question 2
```{r, cache=TRUE}
set.seed(333)
kmeans1 = kmeans(edata_centered, centers = 2)
names(kmeans1)
Vec1 = svd3$v[,1]
cor(Vec1, kmeans1$centers[1,])
```

### Question 3
There are very few samples with more than 2 replicates so the estimates for those values will not be very good.
```{r Q3, cache=TRUE}
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/bodymap_eset.RData")
load(file=con)
close(con)
bm = bodymap.eset
edata = exprs(bm)
pdata_bm=pData(bm)
names(pdata_bm)
table(pdata_bm$num.tech.reps)
Lm1 = lm(edata[1,] ~ as.factor(pdata_bm$num.tech.reps))
names(Lm1)
tidy(Lm1)
par(pch = 19)
plot(edata[1,], col = as.numeric((pdata_bm$num.tech.reps)))
```


### Question 4
```{r Q4}
Lm2 = lm(edata[1,] ~ pdata_bm$age + pdata_bm$gender)
tidy(Lm2)
```
This coefficient means that for each additional year of age, the count goes down by an average of 23.91 for a fixed sex.


### Question 5 and 9
```{r Q5}
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/montpick_eset.RData")
load(file=con)
close(con)
mp = montpick.eset
pdata=pData(mp)
edata=as.data.frame(exprs(mp))
fdata = fData(mp)
edata_tr = log2(edata +1)
library(limma)
mod = model.matrix(~ as.factor(pdata$population))
fit = lm.fit(mod, t(edata))
names(fit)
dim(fit$residuals)
dim(fit$effects)
dim(fit$coefficients)
table(pdata$population, pdata$study) # Question 9
```
The effects are difficult to distinguish because the study variable and population variable are perfectly correlated.     


### Question 6 NOT CORRECT
The model coefficients for all samples for each gene, with the values for each gene stored in the columns of the matrix.     
The definition of effects is given as : n (number of genes) vector of orthogonal single-df effects. The first rank of them correspond to non-aliased coefficients, and are named accordingly.
```{r Q6, fig.align = "center"}
rownames(fit$coefficients)
rownames(fit$effects)[1:4]
par(mfrow = c(1,2))
plot(-fit$effects[1,], col = 2)
plot(fit$coefficients[1,], col = 3)
par(mfrow = c(1,1))
```



### Question 7
```{r Q7}
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/bodymap_eset.RData")
load(file=con)
close(con)
bm = bodymap.eset
edata = exprs(bm)
pdata_bm=pData(bm)
y = pdata_bm$age[!is.na(pdata_bm$age)]
mod = model.matrix(~ y)
lm2 = lmFit(edata[, !is.na(pdata_bm$age)], mod)
names(lm2)
dim(lm2$coefficients)
colnames(lm2$coefficients)
lm2$coefficients[1000,2]
plot(edata[1000,!is.na(pdata_bm$age)] ~ y, col =2)
abline(lm2$coefficients[1000,], col = 3, lwd =2)
```

### Question 8
```{r Q8}
mod1 = model.matrix(~ y + as.factor(pdata_bm$tissue.type[!is.na(pdata_bm$age)]))
mod1 = model.matrix(~ y + as.factor(pdata_bm$tissue.type[!is.na(pdata_bm$age)]))
lm2 = lmFit(edata[, !is.na(pdata_bm$age)], mod1)
dim(mod1)
```


### Question 10
```{r Q10}
set.seed(33353)
edata_tr = log2(edata + 1)
dim(edata_tr)
edata_tr = edata_tr[rowMeans(edata_tr)>=1, ]
dim(edata_tr)
mod1 = model.matrix(~ y, data = pdata_bm[!is.na(pdata_bm$age),])
mod0 = model.matrix(~ 1, data = pdata_bm[!is.na(pdata_bm$age),])
library(sva)
sva1 = sva(dat= edata_tr[,!is.na(pdata_bm$age)], mod = mod1, mod0 = mod0, n.sv = 1)
cor(sva1$sv, y)
summary(lm(sva1$sv ~ pdata_bm$gender[!is.na(pdata_bm$age)]))
summary(lm(sva1$sv ~ pdata_bm$race[!is.na(pdata_bm$age)]))
```



### Session Info
```{r sessionInfo, cache = TRUE}
devtools::session_info()
```

### Last compilation

Last compiled at `r Sys.Date()`.